---
title: "Homework 4"
subtitle: "Prof: Luiz Peternelli"
author: "Marco Guerra Hinostroza"
date: "`r Sys.Date()`"
format:
  html:
    embed-resources: true
    toc: true
    toc-location: left
    number-sections: false
    code-fold: false
    filters:
        - highlight-text
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
library(readODS)
library(dplyr)
library(kableExtra)
```


# Question 1

Probability density function:

$$
f(x) = 
\begin{cases}
1 - x^2 + x^3 \quad ; \quad 0 \le x \le 1 \\
0 \quad ; \quad c.c
\end{cases}
$$
Since the area under the curve is different from 1, it is necessary to transform the density function.

$$
 \int_0^1 k(1-x^2+x^3) \, dx = 1
$$

$$
k \left( \int_0^1 1 \, dx - \int_0^1 x^2 \, dx + \int_0^1 x^3 \, dx \right) = 1
$$

$$
k = \frac{12}{11}
$$

Transformed function:

$$
f(x) = \frac{12}{11} \left(1 - x^2 + x^3 \right) \quad ; \quad 0 \le x \le 1
$$

### Simulation

To optimise the process we can graph the function and define a maximum value (m).
```{r}
curve((12/11)*(1 - x^2 + x^3), from = 0, to = 1)
```

Examining the graph we can take 1.1 as the maximum value.

```{r}
sim_ar <- function(n,m) {
  x <- numeric(n)
  #incluindo um contador para o tamanho da amostra
  i <- 0
  #incluindo um contador para o numero de pares de uniformes geradas
  j <- 0
  
  while (i < n) {
    # 1. Gerar x candidato da distribuição proposta (Uniforme[0, 1])
    xc <- runif(1, min = 0, max = 1)
    
    # 2. Gerar U da Uniforme[0, 1] para a aceitação
    U <- runif(1)
    
    f <- (12/11)*(1 - xc^2 + xc^3)
    
    # 3. Teste de aceitação
    if (U < f/m) {
      x[i+1] <- xc  # Aceita o valor candidato
      i <- i+1  # Incrementa o contador de aceitação
    }
    j <- j+1
  }
  cat("were generated ", j, "pair of uniforms ",fill=T)
  cat("to simulate ", i, " values of the random variable of interest ",fill=T)
  cat("ratio of efficiencx ",round((i/j)*100,2), "% ",fill=T)
  return(x)
}
```

```{r}
x_t <- sim_ar(1000, 1.1)
```

```{r}
hist(x_t, prob = TRUE)
curve((12/11)*(1 - x^2 + x^3), from = 0, to = 1, add = TRUE, col = "blue")
```

# Question 2

```{r, include=FALSE}
dt <- data <- read_ods("tabla.ods", sheet = 1, col_names = TRUE)
```

```{r,echo=FALSE}
colnames(dt) <- c("Z", "Y=0", "Y=1", "Y=0", "Y=1", "Y=0", "Y=1")

kbl(dt) |> 
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left") |> 
  add_header_above(c(" " = 1, "X = 0" = 2, "X = 1" = 2, "X = 2" = 2))
```

### Marginal distribution X
$$
P(X=0) = P(0)= \sum_{y=0}^{2}\sum_{z=0}^{2}p(x,y,z)
$$

$$
P(X=0) = P(0)= \frac{2}{27} + \frac{1}{27} + \frac{5}{27} + \frac{1}{27} = \frac{9}{27}
$$

$$
P(X=1) = P(1)= \sum_{y=0}^{2}\sum_{z=0}^{2}p(x,y,z)
$$

$$
P(X=1) = P(1)= \frac{1}{27} + \frac{5}{27} + \frac{1}{27} + \frac{2}{27} = \frac{9}{27}
$$

$$
P(X=2) = P(2)= \sum_{y=0}^{2}\sum_{z=0}^{2}p(x,y,z)
$$

$$
P(X=2) = P(2)= \frac{3}{27} + \frac{3}{27} + \frac{1}{27} + \frac{2}{27} = \frac{9}{27}
$$

```{r}
x <- 0:3
fx <- c(0, 0.333, 0.333, 0.333)

Fx <- cumsum(fx)
n <- length(x)

plot(x = NA, y = NA, pch = NA, 
     xlim = c(0, max(x)), 
     ylim = c(0, 1),
     xlab = "X",
     ylab = "F(X)",
     main = "Cumulative distribution function")

points(x = x[-n], y = Fx[-1], pch=19)
points(x = x[-1], y = Fx[-1], pch=1)
for(i in 1:(n-1)) points(x=x[i+0:1], y=Fx[c(i,i)+1], type="l")
```

### Conditional distribution (Y|X)

$$
P(Y|X) = \frac{P(Y=y,X=x)}{P(X=x)}=\frac{P(y,x)}{P(x)}
$$

##### For $P(Y|X=0)$

$$P(Y=0|X=0) = \frac{3/27}{9/27} = \frac{1}{3}$$

$$P(Y=1|X=0) = \frac{6/27}{9/27} = \frac{2}{3}$$



##### For $P(Y|X=1)$

$$P(Y=0|X=1) = \frac{6/27}{9/27} = \frac{2}{3}$$

$$P(Y=1|X=1) = \frac{3/27}{9/27} = \frac{1}{3}$$


##### For $P(Y|X=2)$

$$P(Y=0|X=2) = \frac{6/27}{9/27} = \frac{2}{3}$$

$$P(Y=1|X=2) = \frac{3/27}{9/27} = \frac{1}{3}$$

:::: {.columns}

::: {.column width="49%"}
```{r, echo=FALSE}
x <- 0:2
fx <- c(0, 0.333, 0.666)

Fx <- cumsum(fx)
n <- length(x)

plot(x = NA, y = NA, pch = NA, 
     xlim = c(0, max(x)), 
     ylim = c(0, 1),
     xlab = "Y",
     ylab = "F(Y | X = 0)",
     main = "Conditional Cumulative Distribution of Y given X=0")

points(x = x[-n], y = Fx[-1], pch=19)
points(x = x[-1], y = Fx[-1], pch=1)
for(i in 1:(n-1)) points(x=x[i+0:1], y=Fx[c(i,i)+1], type="l")
```
:::

::: {.column width="49%"}
```{r, echo=FALSE}
x <- 0:2
fx <- c(0, 0.666, 0.333)

Fx <- cumsum(fx)
n <- length(x)

plot(x = NA, y = NA, pch = NA, 
     xlim = c(0, max(x)), 
     ylim = c(0, 1),
     xlab = "Y",
     ylab = "F(Y | X = 1)",
     main = "Conditional Cumulative Distribution of Y given X=1")

points(x = x[-n], y = Fx[-1], pch=19)
points(x = x[-1], y = Fx[-1], pch=1)
for(i in 1:(n-1)) points(x=x[i+0:1], y=Fx[c(i,i)+1], type="l")
```
:::

::::

:::: {.columns}

::: {.column width="49%"}
```{r, echo=FALSE}
x <- 0:2
fx <- c(0, 0.666, 0.333)

Fx <- cumsum(fx)
n <- length(x)

plot(x = NA, y = NA, pch = NA, 
     xlim = c(0, max(x)), 
     ylim = c(0, 1),
     xlab = "Y",
     ylab = "F(Y | X = 2)",
     main = "Conditional Cumulative Distribution of Y given X=2")

points(x = x[-n], y = Fx[-1], pch=19)
points(x = x[-1], y = Fx[-1], pch=1)
for(i in 1:(n-1)) points(x=x[i+0:1], y=Fx[c(i,i)+1], type="l")
```
:::

::::

### Conditional distribution (Z|X,Y)

$$
P(Z|X,Y)) = \frac{P(Z=z,X=x,Y=y)}{P(X=x,Y=y)} =\frac{P(z,x,y)}{P(x,y)}
$$

$$

$$

### Simulation
```{r}

```

