---
title: "Linear Regression"
author: "Marco Guerra Hinostroza"
date: "`r Sys.Date()`"
format:
  html:
    embed-resources: true
    toc: true
    toc-location: left
    number-sections: false
    code-fold: false
    filters:
        - highlight-text
editor_options: 
  chunk_output_type: console
---

```{css, echo = FALSE}
.justified {
    text-align: justify;
}
```

```{r setup, include=FALSE}
library(knitr)
library(agridat)
library(tidyverse)
library(tidymodels)
```

# Theoretical questions

### Q1

### Q2

### **Q3. Why is the application of F-statistics in multiple linear regression more appropriate?**

### Q4

### Q5

# Practical questions

### **Q6. Evaluating the association between X and Y**
```{r}
data("reid.grasses")
data <- reid.grasses
```

$$y_i = \beta_0 + \beta_{1}x_i + \epsilon$$
Where $\beta_0$ is the intercept, $\beta_1$ is the effect of the nitrogen level, and $\epsilon$ is the effect of error.

```{r}
model <- lm(drymatter ~ nitro, data = data)
```

Hypothesis testing

Ho : There is no relationship between X and Y ($\beta_1$ $=$ 0).

Ha : There is some relationship between X and Y ($\beta_1$ $\neq$ 0).

We use the t-statistic, t-statistic given by

$$t = \frac{\hat{\beta_1} - 0}{SE(\hat{\beta_1})}$$

:::: {.columns}

::: {.column width="50%"}
$$\hat{\beta_1} = \frac{\sum_{i=1}^{n}(x_i - \bar{x})(y_i - \bar{y})}{\sum_{i=1}^{n}(x_i - \bar{x})^2}$$
:::

::: {.column width="50%"}
$$SE(\hat{\beta_1}) = \sqrt{\frac{\sigma^2}{\sum_{i=1}^{n}(x_i - \bar{x})^2}}$$
:::

::::


```{r}
# Calculate β1 manually using the formula
nitro_mean <- mean(data$nitro)
drymatter_mean <- mean(data$drymatter)
beta1 <- sum((data$nitro - nitro_mean)*(data$drymatter - drymatter_mean)) /
  sum((data$nitro - nitro_mean)^2)

# Calculate σ²
residuals <- model$residuals
n <- length(residuals)  # number of observations
p <- length(coef(model))  # number of parameters (including intercept)
observed_values <- model$model$drymatter
predicted_values <- fitted(model)
sigma2 <- sum((observed_values - predicted_values)^2)/(n - p)

# Calculate SE(β1)
se_beta1 <- sqrt(sigma2 / sum((data$nitro - nitro_mean)^2))

# Calculate t-statistic
t_statistic <- (beta1 - 0) / se_beta1
t_statistic
```

Considering a significance level of 0.001.
```{r}
# Critical t-value for a two-tailed test
t_crt <- qt(1 - 0.001/2, n - 2)
t_crt
```

[Conclusion]{colour="#FFFFFF" bg-colour="#b22222"}\
At a significance level of 0.001 we reject the Ho. Therefore, there is an association between nitrogen level and dry matter.

### **Q7. Estimate the model fit with $R^2$**
To calculate $R^2$, we use the formula

$$R^2 = \frac{TSS - RSS}{TSS} = 1 - \frac{\sum(y_i  - \hat{y}_i)^2}{\sum(y_i - \bar{y})^2}$$
Where $y_i$ is the i-th observed response value, $\hat{y}_i$ is the i-th predicted response value, and $\bar{y}$ is the mean of the observed values.

```{r}
observed_values <- model$model$drymatter
predicted_values <- fitted(model)

# Total Sum of Squares (TSS)
tss <- sum((observed_values - mean(observed_values))^2)

# Residual Sum of Squares (RSS)
rss <- sum((observed_values - predicted_values)^2)

# R^2
r_squared <- 1 - (rss / tss)
r_squared
```

[Summary of results of the fitted model]{color="#FFFFFF" bg-color="#0000FF"}
```{r}
summary(model)
```

### **Q8. LMM**

```{r}
mmodel <- lm(protein ~ nitro + drymatter, data = data)
```

```{r}
summary(mmodel)
```

### **Q9. MLR with discrete variables**

```{r}
mdmodel <- lm(protein ~  nitro + gen + year, data = data)
```

```{r}
summary(mdmodel)
```

```{r}
contrasts(gen)
contrasts(year)
```


### **Q10. MLR with discrete variables**
```{r}

```


# Writing Functions
```{r}

```

